% ******************************* Thesis Dedication ********************************

\begin{acknowledgements} 
The work presented in this thesis was only possible with the support of many people. First and foremost, thank you to Dr. Kellie Morrissey, who has provided guidance and mentored me throughout my undergraduate, master's and PhD. It has been an absolute joy to work with Kellie over the years, and I am genuinely saddened that the collaboration is coming to a close. I hope our work will cross paths again or that myself, you, Jazz, and Daniel Welsh will have another Karo-CHI night. Either will do just fine. Kellie has gone above and beyond as a supervisor, where she has taught me ethical design, academic writing, participatory approaches, and many more that can be found in this thesis. I am deeply appreciative of Kellie's time mentoring me, which has been an invaluable experience that will stick with me forever. Kellie has allowed me to grow and develop my ideas while challenging me to build them further. While I may no longer be in academia, the skills and lessons I have learnt from Kellie have influenced the work I do now, and I will be forever grateful for that.

Thank you to Prof. David Kirk and Prof. Abi Durrant for supporting my ideas over the past few years, who have helped refine my thesis. Dr. Kyle Montague for guidance in the MRes and first year of the PhD. I'd also like to thank all the excellent collaborators I have worked with: Dr Sarah Foley, Dr Rens Brankaert, Dr Jennifer Boger, Dr Amanda Lazar, Dr Madeline Balaam, Dr Gail Kenning, Dr David Unbehaun, Dr Dan Lambton-Howard, and Dr Laura Booi. You have all been incredible to work with and inspired and encouraged me to continue this work over the years. Also, thank you for the humour and support from my amazing Cohort family, who all set out on this weird but wonderful journey of Digital Civics! Oh also, cheers, Rob Anderson, for dealing with me during the write-up year, whether at home or sitting right next to you at work...I'm sorry for all the PhD complaining. 

To Sandra Coulter and Silverline Memories, this research would not have been possible without your kindness, enthusiasm and effort in supporting multiple studies in the thesis. You and the members of Silverline welcomed me and the research with open arms, and I cannot express in writing how grateful I am for all of you. I enjoyed every second of getting to know everyone at Silverline Memories and hearing about their experiences with dementia. Additionally, thank you to the attendees to DemVR; academic researchers who discussed ethics; and the developers, designers and people with dementia who have all helped create this work. Not only have I learnt a lot about dementia through getting to know these people, I also have learned many life lessons that influence who I am today. For that, I am very grateful to everyone who has willingly shared stories, wisdom, guidance and kept me going along this journey.

To all of you, Thank you.
\end{acknowledgements}

