\chapter{Publications arising from thesis}
\label{Publications}
Research presented in this thesis are published in the following peer-reviewed conferences:

\begin{enumerate}

    \item \textbf{James Hodge}, Sarah Foley, Dan Lambton-Howard, Laura Booi, Kyle Montague, Sandra Coulter, David Kirk, and Kellie Morrissey. 2022. Exploring Participants’ Representations and Shifting Sensitivities in a Hackathon for Dementia. ACM Trans. Comput.-Hum. Interact. Just Accepted (November 2022). \href{https://doi.org/10.1145/3571814}{https://doi.org/10.1145/3571814}

    \item \textbf{James Hodge} and Kellie Morrissey. 2020. Sharing a Virtual World with People Living with Dementia. In HCI and Design in the Context of Dementia (pp. 237-253). Springer, Cham. \href{https://doi.org/10.1007/978-3-030-32835-1\_15}{https://doi.org/10.1007/978-3-030-32835-1\_15}

    \item \textbf{James Hodge}, Sarah Foley, Rens Brankaert, Gail Kenning, Amanda Lazar, Jennifer Boger, and Kellie Morrissey. 2020. Relational, Flexible, Everyday: Learning from Ethics in Dementia Research. In Proceedings of the 2020 CHI Conference on Human Factors in Computing Systems (CHI '20). Association for Computing Machinery, New York, NY, USA, 1–16. \href{https://doi.org/10.1145/3313831.3376627}{https://doi.org/10.1145/3313831.3376627}

    \item Rens Brankaert, Gail Kenning, Daniel Welsh, Sarah Foley, \textbf{James Hodge}, and David Unbehaun. 2019. Intersections in HCI, Design and Dementia: Inclusivity in Participatory Approaches. In Companion Publication of the 2019 on Designing Interactive Systems Conference 2019 Companion (DIS '19 Companion). Association for Computing Machinery, New York, NY, USA, 357–360. \href{https://doi.org/10.1145/3301019.3319997}{https://doi.org/10.1145/3301019.3319997}

    \item \textbf{James Hodge}, Kyle Montague, Sandra Hastings, and Kellie Morrissey. 2019. Exploring Media Capture of Meaningful Experiences to Support Families Living with Dementia. In Proceedings of the 2019 CHI Conference on Human Factors in Computing Systems (CHI '19). Association for Computing Machinery, New York, NY, USA, Paper 423, 1–14. \href{https://doi.org/10.1145/3290605.3300653}{https://doi.org/10.1145/3290605.3300653}

    \item \textbf{James Hodge}, Madeline Balaam, Sandra Hastings, and Kellie Morrissey. 2018. Exploring the Design of Tailored Virtual Reality Experiences for People with Dementia. In Proceedings of the 2018 CHI Conference on Human Factors in Computing Systems (CHI '18). Association for Computing Machinery, New York, NY, USA, Paper 514, 1–13. \href{https://doi.org/10.1145/3173574.3174088}{https://doi.org/10.1145/3173574.3174088}

\end{enumerate}
\textbf{Contribution statement:} 
\begin{itemize}
    \item \textbf{Chapter four} - The two studies covered in this chapter were peer-reviewed and published at the CHI Conference on Human Factors in Computing Systems \citep{hodge_exploring_2018,hodge_exploring_2019}. Both papers were co-authored by Dr. Kellie Morrissey and Sandra Hastings, with additional co-authorship from Dr Madeline Balaam for the 2018 paper, and Dr Kyle Montague for the 2019 paper. While this chapter expands on the data that was not used for either publications, I acknowledge that ideas and arguments in the chapter are influenced by the publications. Furthermore, the concluding sections of this chapter act as a set of design decisions that are described in a published book chapter for 'HCI and Design in the Context of Dementia \citep{hodge2020sharing} that I co-authored with Dr Kellie Morrissey who provided feedback on the paper's writing and contribution. 

    \item \textbf{Chapter five} - The study covered in this chapter is currently being peer-reviewed in the TOCHI journal. The paper is co-authored by Dr. Sarah Foley, Dr. Daniel Lambton-Howard, Dr. Kyle Montague, Dr. Laura Booi, Sandra Hastings, Prof. David Kirk, and Dr. Kellie Morrissey. I was responsible for concept, study design, data collection, data analysis and writing of the paper. Co-authors supported the facilitation of the hackathon, and Dr. Kellie Morrissey, Dr. Sarah Foley and Dr. David Kirk provided feedback on the paper's writing and contribution.
    
    \item \textbf{Chapter six} - A version of this chapter was published at CHI'20 \citep{hodge_relational_2020} with collaboration from  Dr. Sarah Foley, Dr. Rens Brankaert, Dr. Gail Kenning, Dr. Amanda Lazar, Dr. Jennifer Boger, and Dr. Kellie Morrissey. In this study, I was responsible for designing the study methodology, organising data collection, data analysis, leading the study, and writing the paper. Several of the authors contributed to conducting the interviews which I analysed. While I envisioned and directed the study, the paper is highly collaborative and it must be acknowledged that many of the authors influenced the ideas and arguments of the paper.

    \item \textbf{Chapter seven} - While the work in chapter seven has yet to be published, Dr. Kellie Morrissey and Prof. David Kirk supported this chapter. I was responsible for concept, study design, data collection, data analysis and writing of the paper. Co-authors supported the final part of the analysis and provided feedback on study design.
\end{itemize}